On the Construction of a Collision-Free Schedule of Beacons in a Wireless Multi-Hop Network

### 2013-10-09 paper.tex
looking for Alessandro s reference

\section{Related Work}
\cite{cano2013iri}

\section{IEEE 802.11s Mesh Beacon Collision Avoidance}
A Mesh Beacon Collision Avoidance (MBCA) is described in \cite{hiertz2007pis}.
The solution consists of sometimes delaying the beacon transmission to check for possible overlapping transmissions among the one-hop neighbours.
It also involves including information about the timing of beacons in beacons.

\section{Assumptions}
We model the network topology as a graph.
In the topology graph, the nodes represent wireless stations and the arcs between two nodes mean that those nodes can communicate with each other.
If a node has two neighbours transmitting simultaneously, it is assumed that the messages are garbled an lost.
A node can only detect and receive the transmissions of its immediate neighbours, and only if there is a single neighbour active.

These are strong assumptions because, in reality, wireless propagation is much more complicated.
The carrier sense range exceeds the transmission range and there is capture effect.
There are a large number of factors that determine whether a transmission is successful or not.
Nevertheless, for tractability, we use the above described graph model.

\section{A Beacon Schedule}
A recurrent solution in wireless networks is the use of beacons for synchronization and signaling.
Beacon messages are very short, which means that the likelihood of collisions is lower.
This beacons can be very useful for network coordination and management.

A schedule of beacons defines special times in which beacons can be transmitted.
Let us assume that beacons should be transmitted every $T_b$ seconds.
Then, it is also necessary to define a schedule length $S$ which is the number of beacons in a schedule.
Each station transmits a beacon once in every schedule in one of the $S$ allowed time instants.
When the schedule is over, it is repeated again.

A beacon includes a list of all the beacons received by the station in the last schedule.
When a station $A$ receives beacons from all its neighbors confirming that they have received the beacon of station $A$, this station can assume that its beacon has not suffered any collision.

@@

\section{The interference graph}
From the topology graph we derive the interference graph.
There is an arc between two nodes the simultaneous transmission of those two nodes results in interference.
Given the interference graph, the problem reduces to map coloring problem.

@@

\section{The construction of a beacon schedule}

@@

\section{Distributed Adjustment of the Schedule Length}
The beacon interval $T_b$ is the same for all the participating stations in the network and is set by the standard or network administrator.
The schedule length should be adjusted according to the node density.

@@

The nodes share this information with their neighbours and then adjusts its own schedule according to the following rules.
\begin{itemize}
\item The schedule is doubled if any of the neighbours report occupation equal to 100\%.
\item The schedule is doubled halved if all the neighbors report occupation higher than 75\%.
\item The schedule is halved if all the neighbors report occupation lower than 25\%.
\end{itemize}

\section{Data Included in the Beacons}
In each beacon, a node includes acknowledgements off all the beacons received in the last schedule, including their position.

@@

\section{The Medium Access Control Protocol}
Each station owns the time that follows its beacon until the next beacon spot.
It can use this time to transmit packets to its neighbours and to request packets from its neighbours using \emph{poll} packets.

@@

>>>>

In this paper, we propose the initial construction of a collision-free schedule of beacons.
This process has to be decentralized as it takes place before the network is operative, which means that the participating stations can exchange data only with their neighbors.
The construction of the schedule is an iterative trial-and-error process that continues until a collision-free schedule of beacons is built.
At that point, it is possible for the nodes to reliable exchange information and transmit data with success guarantees.

\section{The Construction of the Collision-Free Schedule}

\subsection{Assumptions}
We model the network topology as a graph.
In the topology graph, the nodes represent wireless stations and the arcs between two nodes mean that those nodes can communicate with each other.
For illustration purposes, we will use the ring topology in Fig.~\ref{fig:topology}.

If a node has two neighbours transmitting simultaneously, it is assumed that that particular node cannot decode either of the transmissions.
For example, assume that at a given point of time nodes $F$ and $B$ in Fig.~\ref{fig:topology} are active.
Node $A$ will be unable to recover the transmissions.

Note that collisions occur at reception.
The fact that a collision occurs at $A$ when $F$ and $B$ simultaneously transmit does not mean that $E$ or $C$ cannot recover those messages, even though they overlap in time.
A transmission may collide in one receiver and be successful in another one.

A node can only detect and receive the transmissions of its immediate neighbours, and only if there is a single neighbour is active for the whole duration of the transmission.

These graph model represents strong assumptions because, in reality, wireless propagation is much more complicated.
The carrier sense range exceeds the transmission range and there is capture effect.
There are a large number of factors that determine whether a transmission is successful or not.
Nevertheless, for tractability, we use the above described graph model.

\subsection{A Beacon Schedule}
A recurrent solution in wireless networks is the use of broadcast beacons for synchronization and signaling.
Beacon messages are very short, which means that the likelihood of collisions is lower.
This beacons can be very useful for network coordination and management.

A schedule of beacons defines special times in which beacons can be transmitted.
Let us assume that beacons should be transmitted every $T_b$ seconds.
We will name each of the time instants at which a beacon can be transmitted a \emph{spot}.
Each spot can be either empty, successful or collision.
Note that, again, the perception of the spot depends on the position of the station in the topology.
For a given slot, it may happen that one station detects it as empty, another station as successful and another station as collision.

Then, it is also necessary to define a schedule length $S$ which is the number of spots.
Each station transmits a beacon once in every schedule in one of the $S$ allowed spots.
When the schedule is over, it is repeated again.
The length of the schedule in time units is simply $T_s=S*T_b$.

An example of a beacon schedule is represented in Fig.~\ref{fig:timeline}.
The axis represents time and each of the shaded blocks represents an spot in which beacon transmissions occur.
The dotted box is a spot in which no transmission occur.

Note that the starting and ending position of the schedule is only of local significance. 
We take the convention that each node considers the initial position of the schedule the one in which it transmits, and is numbered as zero.

A beacon includes acknowledgement information for all the beacons received by the station in the last schedule.
When a station $A$ receives beacons from all its neighbors confirming that they have received the beacon of station $A$, this station can assume that its beacon has not suffered any collision.

@@

\subsection{The Interference Graph}
From the topology graph we derive the interference graph.
In the interference graph, there is an arc between two nodes the simultaneous transmission of those two nodes results in interference.
Given the interference graph, the problem finding the collision-free schedule of beacons reduces to map coloring problem.

@@

\subsection{The construction of a beacon schedule}

@@

\subsection{Distributed Adjustment of the Schedule Length}
The beacon interval $T_b$ is the same for all the participating stations in the network and is set by the standard or network administrator.
The schedule length should be adjusted according to the node density.

@@

The nodes share this information with their neighbours and then adjusts its own schedule according to the following rules.
\begin{itemize}
\item The schedule is doubled if any of the neighbours report occupation equal to 100\%.
\item The schedule is doubled if all the neighbors report occupation higher than 75\%.
\item The schedule is halved if all the neighbors report occupation lower than 25\%.
\end{itemize}

\subsection{Data Included in the Beacons}
In each beacon, a node includes acknowledgements off all the beacons received in the last schedule, including their position.

@@

\subsection{The Medium Access Control Protocol}
Each station owns the time that follows its beacon until the next beacon spot.
It can use this time to transmit packets to its neighbours and to request packets from its neighbours using \emph{poll} packets.

@@

\section{Related Work}
An initial step in the direction we are exploring in this later was taken in \cite{cano2013iri}.
In that paper, the construction of a collision-free schedule is studied for a sensor network.
That paper focused on a simple scenario in which all nodes could hear each other transmissions.
Another difference between that paper and the present letter is that here we not aim to compute a schedule length which is similar or equal to the participating nodes.
In the present letter the schedule is independently computed by each node and therefore it is possible that different nodes use different schedule lengths.

\subsection{IEEE 802.11s Mesh Beacon Collision Avoidance}
A Mesh Beacon Collision Avoidance (MBCA) is described in \cite{hiertz2007pis}.
The solution consists of sometimes delaying the beacon transmission to check for possible overlapping transmissions among the one-hop neighbours.
It also involves including information about the timing of beacons in beacons.
